---
title: "insideMPA"
author: "Xuelei_Zhang"
date: '2022-12-02'
output: html_document
---


Table 2.3 calculate locations inside and outside MPAs 

Summary:cutoff g=0.5143

      Auckland                                Campbell Is                            
      inside/outside   search/transit        inside/outside   search/transit     
2022: 628/1737         295/333                  0/967            NA
2021: 430/820          292/138                 31/1219           13/18
2020: 268/332          195/73                   6/594            0/6 
           
     subantarctic IMMA                       Rakirua/Stewart Island       
2022: 500/467          303/197                  17/950        0/17
2021: 837/413          460/377                  13/1237       0/13
2020: 378/222          269/109                  0/600          NA



```{r inside=F}

library(dplyr)
library(ggplot2)
library(tidyverse)
library(sf)
library(rnaturalearth)
library(rgdal)
library(viridis)



```


```{r}

MPA_whole<-read_sf("D:/auckland/nzsrw/maps/maps/proposal/marinereserves/marine-reserves.shp")


aucklandMPA <- MPA_whole %>% subset  (OBJECTID=="67220")
campbellMPA <-  MPA_whole %>% subset  (OBJECTID=="163562")

remove(MPA_whole)
```

Auckland Island
      Inside/outside   Search/transit 
2022: 628/1737         295/333
2021: 582/1084        367/215
2020: 258/336          177/81
2009: 309/169          213/96

Campbell Island:
      Inside/outside   Search/transit 
2022:  31/2334          26/5
2021:  29/1637          18/11
2020:  6/588            0/6
2009:  0/478            0/0

```{r inside auckland is mpa 2009}

load("D:/auckland/nzsrw/chapter1 07102023/2009/mpm_ssm_12h_2009.Rdata")

sf_09 <- st_as_sf(mpm_ssm_12h, coords = c("lon", "lat"), 
    crs = 4326) #%>% 
    #st_transform(crs= 3851)

aucklandMPA<-st_transform(x=aucklandMPA,crs=4326)

inside_akl_mpa_09 <- st_intersects(x=sf_09,y=aucklandMPA,sparse = F)
table(inside_akl_mpa_09)

# 309 points inside MPA 
# 319 PTS outside

# keep points only inside MPA 
sf_09 <- sf_09%>%
        dplyr::mutate(sf_09,inside_akl_mpa=inside_akl_mpa_09)%>%
        filter(inside_akl_mpa_09=="TRUE")
        

# create cutoff = 0.28831 search/transit=295/333

sf_09$state <- cut(sf_09$g,breaks=c(0, 0.28831,1),labels=c("search","transit"),include.lowest=TRUE)

table(sf_09$state)


```

```{r inside Auckland IS MPA 2022}

 load("D:/auckland/nzsrw/chapter1 07102023/2022/ssm_mpm_22.Rdata")

sf_22 <- st_as_sf(ssm_mpm_22, coords = c("lon", "lat"), 
    crs = 4326) #%>% 
    #st_transform(crs= 3851)

aucklandMPA<-st_transform(x=aucklandMPA,crs=4326)

inside_akl_mpa_22 <- st_intersects(x=sf_22,y=aucklandMPA,sparse = F)
table(inside_akl_mpa_22)

# 628 points inside MPA 
# 1737 PTS outside

# keep points only inside MPA 
sf_22 <- sf_22%>%
        dplyr::mutate(sf_22,inside_akl_mpa=inside_akl_mpa_22)%>%
        filter(inside_akl_mpa_22=="TRUE")
        

# create cutoff = 0.5143 search/transit=295/333

sf_22$state <- cut(sf_22$g,breaks=c(0, 0.5143,1),labels=c("search","transit"),include.lowest=TRUE)

table(sf_22$state)


sf::sf_use_s2(FALSE)

ggplot()+
  geom_sf(data=sf_22,size=1.5,aes(col=as.factor(state)))+
  geom_sf(data=aucklandMPA,fill=NA)+
  theme_bw()+
  ggtitle("akl 2022")

```

```{r inside campbell IS MPA 2022}

load("D:/auckland/nzsrw/chapter1 07102023/2022/ssm_mpm_22.Rdata")

sf_22 <- st_as_sf(ssm_mpm_22, coords = c("lon", "lat"), 
    crs = 4326) #%>% 
    #st_transform(crs= 3851)

campbellMPA<-st_transform(x=campbellMPA,crs=4326)

inside_cpb_mpa_22 <- st_intersects(x=sf_22,y=campbellMPA,sparse = F)
table(inside_cpb_mpa_22)

# 31 points inside MPA 
# 2334 pts outside MPA 

sf_22 <- sf_22%>%
        dplyr::mutate(sf_22,inside_cpb_mpa=inside_cpb_mpa_22)%>%
        filter(inside_cpb_mpa_22=="TRUE")
        

# create cutoff = 0.5143 

#search/transit=26/5

sf_22$state <- cut(sf_22$g,breaks=c(0,0.5143,1),labels=c("search","transit"),include.lowest=TRUE)

table(sf_22$state)

```

```{r inside Auckland IS MPA 2021}

load("D:/auckland/nzsrw/chapter1 07102023/2021/ssm_mpm_21.Rdata")

sf_21 <- st_as_sf(ssm_mpm_21, coords = c("lon", "lat"), 
    crs = 4326) #%>% 
    #st_transform(crs= 3851)

aucklandMPA<-st_transform(x=aucklandMPA,crs=4326)

inside_akl_mpa_21 <- st_intersects(x=sf_21,y=aucklandMPA,sparse = F)
table(inside_akl_mpa_21)

# 1666 inside the eez 
# 582 points inside MPA 
# 1084 PTS outside

# keep points only inside eez 
sf_21 <- sf_21%>%
        dplyr::mutate(sf_21,inside_akl_mpa=inside_akl_mpa_21)%>%
        filter(inside_akl_mpa_21=="TRUE")
        

# create cutoff = 0.5143
# search/transit= 367/215

sf_21$state <- cut(sf_21$g,breaks=c(0,0.5143,1),labels=c("search","transit"),include.lowest=TRUE)

table(sf_21$state)


sf::sf_use_s2(FALSE)

ggplot()+
  geom_sf(data=sf_21,size=1.5,aes(col=as.factor(state)))+
  geom_sf(data=aucklandMPA,fill=NA)+
  theme_bw()+
  ggtitle("akl 2021")

```

```{r inside campbell IS MPA 2021}

load("D:/auckland/nzsrw/chapter1 07102023/2021/ssm_mpm_21.Rdata")

sf_21 <- st_as_sf(ssm_mpm_21, coords = c("lon", "lat"), 
    crs = 4326) #%>% 
    #st_transform(crs= 3851)

campbellMPA<-st_transform(x=campbellMPA,crs=4326)

inside_cpb_mpa_21 <- st_intersects(x=sf_21,y=campbellMPA,sparse = F)
table(inside_cpb_mpa_21)

# 29 points inside MPA 
# 1637 PTS outside

# keep points only inside eez 
sf_21 <- sf_21%>%
        dplyr::mutate(sf_21,inside_cpb_mpa=inside_cpb_mpa_21)%>%
        filter(inside_cpb_mpa_21=="TRUE")
        

# create cutoff = 0.5143 search/transit=18/11

sf_21$state <- cut(sf_21$g,breaks=c(0,0.5143,1),labels=c("search","transit"),include.lowest=TRUE)

table(sf_21$state)


sf::sf_use_s2(FALSE)

ggplot()+
  geom_sf(data=sf_21,size=1.5,aes(col=as.factor(state)))+
  geom_sf(data=campbellMPA,fill=NA)+
  theme_bw()+
  ggtitle("cpb 2021")

```

```{r inside Auckland IS MPA 2020}

load("D:/auckland/nzsrw/chapter1 07102023/2020/ssm_mpm_20.Rdata")

sf_20 <- st_as_sf(ssm_mpm_20, coords = c("lon", "lat"), 
    crs = 4326) #%>% 
    #st_transform(crs= 4326)

aucklandMPA<-st_transform(x=aucklandMPA,crs=4326)

inside_akl_mpa_20 <- st_intersects(x=sf_20,y=aucklandMPA,sparse = F)
table(inside_akl_mpa_20)

# 594 inside the eez
# 258 points inside MPA 
# 336 PTS outside

# keep points only inside eez 
sf_20 <- sf_20%>%
        dplyr::mutate(sf_20,inside_akl_mpa=inside_akl_mpa_20)%>%
        filter(inside_akl_mpa_20=="TRUE")
        

# create cutoff = 0.5143 search/transit= 177/81

sf_20$state <- cut(sf_20$g,breaks=c(0,0.5143,1),labels=c("search","transit"),include.lowest=TRUE)

table(sf_20$state)


sf::sf_use_s2(FALSE)

ggplot()+
  geom_sf(data=sf_20,size=1.5,aes(col=as.factor(state)))+
  geom_sf(data=aucklandMPA,fill=NA)+
  theme_bw()+
  ggtitle("akl 2020")

```

```{r inside campbell IS MPA 2020}

load("D:/auckland/nzsrw/chapter1 07102023/2020/ssm_mpm_20.Rdata")

sf_20 <- st_as_sf(ssm_mpm_20, coords = c("lon", "lat"), 
    crs=4326) #%>% 
    #st_transform(crs=9191)

campbellMPA<-st_transform(x=campbellMPA,crs=4326)

inside_cpb_mpa_20 <- st_intersects(x=sf_20,y=campbellMPA,sparse = F)
table(inside_cpb_mpa_20 )

# 594 pts outside campbell is MPA
# 6 inside 
# 588 outside MPA

sf_20 <- sf_20%>%
        dplyr::mutate(sf_20,inside_cpb_mpa_20=inside_cpb_mpa_20)%>%
        filter(inside_cpb_mpa_20=="TRUE")
        
# create cutoff = 0.5143 search/transit=0/6

sf_20$state <- cut(sf_20$g,breaks=c(0,0.5143,1),labels=c("search","transit"),include.lowest=TRUE)

table(sf_20$state)

ggplot()+
  geom_sf(data=sf_20,size=1.5,aes(col=as.factor(state)))+
  geom_sf(data=campbellMPA,fill=NA)+
  theme_bw()

```


# 2022: inside/outside 1653/712 search/transit = 1087/566
# 2021: 1056/610;470/586
# 2020: 377/217; 261/116
# 2009: 367/111; 237/130 cutoff=0.28831; substantially lower than other years


```{r New Zealand Subantarctic Islands IMMA 2022 2021 2020}


IMMA <- readOGR("D:/auckland/nzsrw/maps/maps/proposal/imma/iucn-imma.shp")
IMMA<-st_as_sf(IMMA,crs=4326)

# subset subantarctic IMMA and Rakiura Stewart Island IMMA
subIMMA<-IMMA%>% subset(Ident.Code == "07EXSOOC182020IMMA")

stewIMMA <-IMMA%>% subset( Ident.Code =="21ANSEIO202020IMMA")

remove(IMMA)

load("D:/auckland/nzsrw/chapter1 07102023/2022/ssm_mpm_22.Rdata")

sf_09 <- st_as_sf(mpm_ssm_12h, coords = c("lon", "lat"), 
    crs = 4326) #%>% 
    #st_transform(crs= 3851)

inside_subIMMA_09 <- st_intersects(x=sf_09,y=subIMMA,sparse = F)
table(inside_subIMMA_09)

# 2022: inside/outside 1653/712 search/transit = 1087/566
# 2021: 1056/610;470/586
# 2020: 377/217; 261/116
# 2009: 367/111; 237/130

# keep points only inside sub IMMA 
sf_09<- sf_09%>%
        dplyr::mutate(sf_09,inside_subIMMA_09=inside_subIMMA_09)%>%
        filter(inside_subIMMA_09=="TRUE")

sf_09$state <- cut(sf_09$g,breaks=c(0,0.28831,1),labels=c("search","transit"),include.lowest=TRUE)

table(sf_09$state) 

sf::sf_use_s2(FALSE)

ggplot()+
  geom_sf(data=sf_22,size=1,alpha=0.5,aes(col=as.factor(state)))+
  geom_sf(data=subIMMA,fill=NA)+
  theme_bw()+
  ggtitle("subIMMA 2022")


```

# inside/outsie  0/478   search/transit = 0/0 2009 
# inside/outside 0/594   search/transit = 0/0 2020
# inside/outside 16/1650 search/transit = 0/16 2021
# inside/outside 15/2350 search/transit=0/15  2022


```{r}

load("D:/auckland/nzsrw/chapter1 07102023/2022/ssm_mpm_22.Rdata")
load("D:/auckland/nzsrw/chapter1 07102023/2021/ssm_mpm_21.Rdata")
load("D:/auckland/nzsrw/chapter1 07102023/2020/ssm_mpm_20.Rdata")
load("D:/auckland/nzsrw/chapter1 07102023/2009/ssm_mpm_12h_names_2009.Rdata")

sf_22 <- st_as_sf(ssm_mpm_22, coords = c("lon", "lat"), 
    crs = 4326) #%>% 
    #st_transform(crs= 3851)

sf_21 <- st_as_sf(ssm_mpm_21, coords = c("lon", "lat"), 
    crs = 4326) #%>% 
    #st_transform(crs= 3851)


inside_stewIMMA_21 <- st_intersects(x=sf_21,y=stewIMMA,sparse = F)
table(inside_stewIMMA_21)


# inside/outsie  0/478   search/transit = 0/0 2009 
# inside/outside 0/594   search/transit = 0/0 2020
# inside/outside 16/1650 search/transit = 0/16 2021
# inside/outside 15/2350 search/transit=0/15  2022

# keep points only inside sub IMMA 
sf_21 <- sf_21%>%
        dplyr::mutate(sf_21,inside_stewIMMA_21=inside_stewIMMA_21)%>%
        filter(inside_stewIMMA_21=="TRUE")

sf_21 <- sf_21[-17]


sf_21$state <- cut(sf_21$g,breaks=c(0,0.52,1),labels=c("search","transit"),include.lowest=TRUE)

table(sf_21$state)

sf::sf_use_s2(FALSE)

ggplot()+
  geom_sf(data=sf_22,size=1,alpha=0.5,aes(col=as.factor(state)))+
  geom_sf(data=subIMMA,fill=NA)+
  theme_bw()+
  ggtitle("subIMMA 2022")

save(sf_21, file = "G:/auckland/nz_right_whale/maps/analysis/data processing/new/insideMPA/sf.21.stew.Rdata")

```